### 2024-07-17 thesis/ms-fletcher-porter.tex
s/Fluid power/Mechatronics/ in the abstract
Add final changes suggested by Jari
It's done!!

%% English abstract.
%% All the information required in the abstract (your name, thesis title, etc.)
%% is used as specified above.
%% Specify keywords
%%
%% Kaikki tiivistelmässä tarvittava tieto (nimesi, työnnimi, jne.) käytetään
%% niin kuin se on yllä määritelty.
%% Avainsanat
%%
\keywords{Mechatronics,Control Engineering,Software Tooling,Stewart Platform,Pneumatic actuator}
%% Abstract text
\begin{abstractpage}[english]
Fluid power systems can be enormously complex. The chiefest complexity is in their control components due to the deep knowledge required for effective development, the need to interface with myriad heterogeneous components, and the opacity of its operation. In the IT sector, a tech stack is a known discretization of function within a system, typically used to categorize what software tools do. This enables disparate parts of a system to be developed separately. Inspired by IT tech stacks, this paper presents a model, the mechatronic tech stack, to mitigate this complexity by siloing control system behavior into discrete functions and classifying what tools can be used to meet them. The model is demonstrated on three case studies of diverse application. The theory is promising, but there are challenges in defining the most useful abstractions and interfaces.
\end{abstractpage}

@@

In order to study mechatronics systems, it is necessary to design and develop physical research test beds. This is especially the case in mechatronic systems, as often they have component whose high-fidelity modeling is challenging and computationally expensive. When doing research with hardware systems, before a single research question can be asked of the hardware, researchers must assemble mechanical components, hook up power electronics, and route hoses for the hydraulic and pneumatic systems. Then the most complex of them all, the control system, must be added. Its tangle of spindly wires reaches all throughout the machine, meeting at the computer, the single most complex component of any fluid power system. The requirements placed on it are immense. In high-power systems, safety invariants must be held up perfectly lest disaster strike, and close monitoring is necessary due to the low stiffness of fluid systems compared to mechanical ones.

@@

Another iconic design was the mechanical governor of the seventeenth century that regulated steam engine speeds by exploiting centrifugal forces to raise and lower masses to impede changes in engine speed.  This was another proportional controller.  It was described formally by Maxwell in 1868 where he wrote about its dynamics and stability, noting that the primary stability criterion is that all the roots of the systems dynamics must have negative real parts~\cite{Maxwell1868}.  Today we understand this as the Nyquist Criterion, but it was known 60 years before Nyquist wrote about it.

@@

Typically control and measurement signals consisted of $\qtyrange[range-units=single,range-phrase=..]{4}{20}{\milli\ampere}$ analog signals and $\qty{24}{\volt}$ digital signals.  This meant that every single sensor and actuator needed at least two wires, one for power and one for ground, going from it to the control box.  For large-scale controllers in factories, this was expensive and untenable to maintain~\cite{Thomesse2005}.  To solve this, practitioners developed fieldbus networks to handle their input and output.  The earliest was CAMAC which was used in the 1960s for nuclear instrumentation~\cite{Thomesse2005}.  Fieldbus networks provided benefits additional to wire cleanup.  They also enable distributing control information to many machines at once on a single network, monitoring the state of the factory floor from the design or business sections of the factory, and hierarchical control, such as described by the US National Bureau of Standards AMRF control hierarchy, where more abstract elements of an industrial plants command more concrete ones.  In particular, it describes a facility, consisting of production management and manufacturing engineering, commanding the shop, or factory floor, which in turn commands individual manufacturing cells, which then command the workstations within, and then at those workstations, operators command equipment~\cite{McLean1987}.  This all can be seen in figure~\ref{fig:amrf_control_hierarchy}.  The control tooling discussed in this thesis is concerned with the business at the Equipment and Workstation level.

@@

This paper builds on this history to improve the process of commissioning control systems for mechatronic systems by reasoning about how to abstract the construction of behavior and components in computer control systems into layers, synthesizing a novel {\it mechatronic tech stack\/} theory, not unlike the AMRF control hierarchy. It will demonstrate in three case studies how this theory can be applied in mechatronic systems. The tools used in these cases will be categorized into layers of the mechatronic tech stack.  The analysis shows that, while different, these cases all fit into the same structure.

@@

This section constructs the backbone of this paper, namely the novel theory of a mechatronic tech stack. First, there will be a review of tech stacks and {\it abstraction layers\/} as they've been applied in the IT sector. Then there will be a proposal of a mechatronic tech stack determined by a standards-based set of principles.

@@

Tools in each of these layers are myriad and to highlight that, proceeding in figure~\ref{fig:web_stack} is a list of a small sample. Front-end tools like jQuery~\cite{jQuery} have been proven from many years of use, React~\cite{React} is very powerful and can be used in non-web contexts, and even plain HTML without an additional framework can be good on account of its simplicity. Back-end tools can allow a developer to select their favorite programming languages like Ruby or Python, or one that is specialized for web back-ends like PHP, or even JavaScript like the front end uses. Servers are often integrated into the back-end framework, but if not, specialized servers like Apache~\cite{ApacheServer} or nginx~\cite{nginx} are popular. Databases are dominated by reliable and proven relational, SQL-based offerings like MySQL or PostgreSQL, but alternatives like NoSQL or Redis~\cite{redis} have become popular in the past years.

@@

The model itself consists of layered abstractions over physical communication media, as seen in figure~\ref{fig:osi_model}. The lowest OSI layer, Physical, concerns itself with the physical phenomena of communication. In the case of CAN, this means specifying voltage levels, terminating resistors, twisted pairs, grounding rules, and more. The highest level, Application, provides networking facilities without need for understanding \textit{how} the network connection is made. The Application layer includes, for example, identifying communication partners and negotiating security protocols. Well-known protocols like HTTP, SSH, and FTP exist in the Application layer. The intermediate layers provide increasing levels of abstraction over the physical layer, defining things like the structure of data packets, how to route messages, how to route them efficiently, and more.

@@

When these tokens are combined into phrases, that's syntax. Syntax is the structure into which meaning is applied. The \verb|printf| example is syntactically parsed by C compilers as in figure~\ref{fig:printf_syntax}. The syntax tells that the postfix expression represented by the identifier \verb|printf| has the argument list of \verb|"I have %u cows\n"| (a string literal) and \verb|num_cows| (an identifier). What any of this means is the domain of semantics.

@@

This is enough to propose a mechatronic tech stack. The mechatronic stack as outlined sits on top of signaling with sensors and transducers. Those are interfaced by IO systems that connect to computer systems which run middleware upon which a control program depends. The design is so that functionality in the control layer should be preserved if components in the lower layers are changed, perhaps with some implementation effort, as was the case for IP over CAN.  This is illustrated as a stack diagram in figure~\ref{fig:mechatronic_tech_stack}.

@@

Following will be three case studies providing concrete illustrations of how this mechatronic tech stack model can be applied. These will mostly be descriptive case studies that apply this theory for retrospective research, but one case uses this model prescriptively.

@@

Dolores is a full-scale test bed in the Aalto University Fluid Power Laboratory, seen in figure~\ref{fig:dolores0} to study the application of electric hydrostatic actuation in an articulated-steering linkage for a front-loading mining vehicle~\cite{Hermansson2021, Helduser1999, Dudzinski1989}. Its control system commissioning was completed in 2021, but in 2023, an industry partner in the experiment requested hardware changes that demanded a substantial re-architecture of the control system. This case study will discuss the mechatronic tech stacks behind the old and the new systems.

@@

Together, the stack for the old system was CANopen-based IO with a DAQ supplementing simple analog and digital signals, a workstation computer, Windows middleware with drivers and libraries to support communication with the CAN bus and DAQ, and a controller based on Matlab and Simulink. In the new system, the stack was a mix of EtherCAT and CANopen IO, a Beckhoff Industrial PC in concert with a workstation, PLC middleware with an application to initialize the CAN network, and a controller written in PLC languages with TwinCAT.\ This is illustrated in figure~\ref{fig:mechatronic_tech_stack_dolores}.

@@

A research group at Okayama University of Science has, over the past years, developed a novel soft, pneumatic actuator that extends when supplied with pressure called an extension-type flexible pneumatic actuator (EFPA)~\cite{Noritsugu2005}. This section will study the tech stacks for three applications of this actuator, a six-legged mobile robot for providing the elderly with an interactive physical activity~\cite{Hase2020}, a pipe-traversal robot for inspecting and repairing water pipe networks~\cite{Shinohara2020}, and a lightweight servo valve based on bending hoses to support EFPA applications~\cite{Kobayashi2020}. All of these can be seen in figure~\ref{fig:soft_pneumatic_actuators_demo}. Note that because this is a case study on work I did not participate in, some details are not documented and must be inferred.

@@

The system is a chair atop a Stewart platform whose linear actuators are single-acting pneumatic-muscle cylinders, seen in figure~\ref{fig:stewart_platform_demo}~\cite{Stewart1965, Caldwell1995}. The valve manifolds that were already made for this had one valve to connect the pneumatic muscle to an external pressure source and another to vent the muscle to atmosphere. When both valves are closed, the cylinder brakes. There are six instances of this circuit, connected in series to a common pressure source. Each cylinder also has a position sensor exposed as a voltage signal from a potentiometer.

@@

For the control layer, I wrote the code in the Sublime Text text editor~\cite{SublimeText} and configured it to automatically format the source code according to standards for the Rust language using the \verb|rustfmt| utility~\cite{rustfmt}.  Other text editors like Emacs, Vim, or VSCodium would have also been good choices, but Sublime is familiar.  This was done on a Windows workstation, but it was compiled and run on the Raspberry Pi Linux control computer.  To achieve this, I checked the changes on the workstation into a \verb|git| version control repository and pushed them to the control computer\footnote{A small issue that comes up is that \verb|git| generally doesn't allow users to push changes to a branch of a repository that the local user has checked out, which is the case here~\cite[§receive.denyCurrentBranch]{GitConfig}.  This is to prevent data loss.  To enable this, \verb|git| can be configured to update the local branch only if there is no possibility of data loss by executing \verb|git config receive.denyCurrentBranch updateInstead|.}.  Then, though a remote terminal, I executed commands to compile and run the program.  This became quite annoying with how many steps there are, so I use the utility \verb|cargo-watch| to automatically execute the compile and run commands on the control computer when there are changes\footnote{The command was similar to \verb|cargo watch -x 'run -- eth0'|, which compiles and executes the program with the argument \verb|eth0|, which is the network interface the EtherCAT connection is on.}~\cite{CargoWatch}.  This still is somewhat tedious and so requires more development.  It also would have been mitigated by developing in some kind of Linux environment or by using a containerized environment like would be provided by Docker or Vagrant, which would have allowed running some tests on the development machine~\cite{DockerDev,VagrantDev}.  Most of the middleware was prepared with the Raspberry Pi OS image that was installed while commissioning the control computer, but there was some additional configuration, e.g.\ to install \verb|git|, that was done with the OS's \verb|apt| package management utility~\cite{apt}.  As well, library dependencies for the control program like Ethercrab were managed with the Rust \verb|cargo| utility~\cite{Cargo}.  The computer, once commissioned, did not require any development.  The IO layer's development was aided by a four-channel oscilloscope and a multimeter, seen in figure~\ref{fig:stewart_platform_demo}, to validate electrical connections and ensure signal were propagating as expected.  It was also interesting to observe the timing jitter of the program, which was generally very small, on the order of microseconds.

@@

The Solid Mechanics Laboratory at Aalto University conducts experiments to identify the properties of engineering materials under load particularly studying phenomena ranging from the micron to the meter scale.  As is typical for laboratories studying solid mechanics, they do this by subjecting test samples to a load schedule using a servo-controlled hydraulic cylinder as seen in figure~\ref{fig:solid_mechanics_test_rig}.  They have several test beds, each with an instance of a common control system.  This had been controlled by a workstation-commanded DAQ that communicated with a motor controller and a strain gauge amplifier, but new experiments and the desire for faster, better research cycles called for improvements to the control system.  The following discussion is based heavily on an interview I took with the lab's head Dr.~Pauli Lehto.

@@

The control layer was built, like the lower layers, with NI tooling.  All of it was built with LabView, with the CPU-based controls using the LabView Real-Time Module~\cite{NIRealTimeModule} and the FPGA controls built with the LabView FPGA Module~\cite{NIFPGAModule}.  It implemented a PID controller, though the derivative term was in practice usually omitted~\cite{Schad2011,Ali2010}.  To monitor the condition of long-running tests and for data acquisition, there are programs running on a desktop workstation, seen in figure~\ref{fig:solid_mechanics_control_monitor}, that observe the state of the control system.  They also developed a custom scripting language that runs on the workstation to define the high-level testing schedule, separate from the low-level control logic.

@@

The Aalto Solid Mechanics Laboratory's new control system provides a compelling case study on how control components may be selected when there are requirements placed on the control system beyond easy development.  It led to a need for high-speed IO and computation and a computer with multiple computation units with a CPU and an FPGA.\  Its tech stack is illustrated in figure~\ref{fig:mechatronic_tech_stack_solid_mechanics}.  The system as constructed has been able to meet the performance requirements and execute useful tests.  Development is on-going.

@@

This paper has introduced a theory of a mechatronic tech stack which describes function and tooling for mechatronic systems, namely as discrete layers of control, middleware, computer, and IO.\  This theory is motivated by analogy to the web stack, used in IT software development, and to the OSI networking model, used to describe networked computer systems.  The OSI model also provided a means of discriminating the layers that make up the model, importantly by finding contrasts in the morphology, syntax, and semantics of the behavior of components of the control system.

@@

This theory can also be extended.  Throughout this work, it has been called a \textit{mechatronic} tech stack, but truly it's a control system tech stack.  To make it a true mechatronic tech stack, it should include tooling for mechanical and electronic systems.  A mechanical section of the stack would grow from the bottom of the stack shown in figure~\ref{fig:mechatronic_tech_stack}, not colliding with the current stack, but an electrical component would be more complicated, since it is needed by both the mechanical stack for transducers and sensors, and by the control stack since the computer \textit{is} an electrical component, and the IO layer often needs some electrical finesse, as noted in §\ref{sec:stewart_platform}.  It could also be extended in the software realm; the production of graphical monitoring systems and tele-operation programs generally require substantial non-control software which should be described by different layers, perhaps forking away from the middleware layer to have a monitoring layer branch.

@@

This paper presented three case studies to demonstrate the utility of the novel theory of a mechatronic tech stack. These cases were diverse in their applications and methods and by viewing them through the lens of the tech stack, new knowledge of how mechatronic systems can be commissioned was gained. Also, by studying several stacks, I have formed a pool of tools to draw from. The mechatronic tech stack can be re-rendered to reflect this, as here in figure~\ref{fig:mechatronic_tech_stack_filled}.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{assets/mechatronic_tech_stack_filled}
  \caption{The mechatronic tech stack for all of the case studies in one, with tools for each project highlighted in different colors.  Not only does the mechatronic tech stack present layers of abstraction, it also provides an idea of what tools to choose, similar to the web stack in figure~\ref{fig:web_stack}.}\label{fig:mechatronic_tech_stack_filled}
\end{figure}

>>>>

%% English abstract.
%% All the information required in the abstract (your name, thesis title, etc.)
%% is used as specified above.
%% Specify keywords
%%
%% Kaikki tiivistelmässä tarvittava tieto (nimesi, työnnimi, jne.) käytetään
%% niin kuin se on yllä määritelty.
%% Avainsanat
%%
\keywords{Mechatronics,Control Engineering,Software Tooling,Stewart Platform,Pneumatic actuator}
%% Abstract text
\begin{abstractpage}[english]
Mechatronic systems can be enormously complex. The chiefest complexity is in their control components due to the deep knowledge required for effective development, the need to interface with myriad heterogeneous components, and the opacity of its operation. In the IT sector, a tech stack is a known discretization of function within a system, typically used to categorize types of software tools and the functions they perform. This enables disparate parts of a system to be developed separately with the confidence that other components will or should provide expected interfaces. Inspired by IT tech stacks, this thesis presents a model, the mechatronic tech stack, to mitigate this complexity by siloing control system behavior into discrete functions, classifying what tools can be used to meet them, and how development can be guided by this theory. The model is demonstrated on four case studies related to control upgrades on a fluid power test bench, development of soft robots, the development of a pneumatic muscle Stewart platform, and controllers for a solid mechanics lab. The theory is promising and there is a clear direction to progress, but there are challenges in defining the most useful abstractions and interfaces.
\end{abstractpage}

@@

In order to study mechatronics systems, it is necessary to design and develop physical research test beds. This is especially the case in mechatronic systems, as often they have component whose high-fidelity modeling is challenging and computationally expensive. When doing research with hardware systems, before a single research question can be asked of the hardware, researchers must assemble mechanical components, hook up power electronics, and route hoses for the hydraulic and pneumatic systems. Then the most complex of them all, the control system, must be added. Its tangle of spindly wires reaches all throughout the machine, meeting at the computer, the single most complex component of any mechatronic system. The requirements placed on it are immense. In high-power systems, safety invariants must be held up perfectly lest disaster strike, and close monitoring is necessary due to the low stiffness of fluid systems compared to mechanical ones.

@@

Another iconic design was the mechanical governor of the eighteenth century that regulated steam engine speeds by exploiting centrifugal forces to raise and lower masses to impede changes in engine speed.  This was another proportional controller.  It was described formally by Maxwell in 1868 where he wrote about its dynamics and stability, noting that the primary stability criterion is that all the roots of the systems dynamics must have negative real parts~\cite{Maxwell1868}.  Today we understand this as the Nyquist Criterion, but it was known 60 years before Nyquist wrote about it.

@@

Typically control and measurement signals consisted of $\qtyrange[range-units=single,range-phrase=..]{4}{20}{\milli\ampere}$ analog signals and $\qty{24}{\volt}$ digital signals.  This meant that every single sensor and actuator needed at least two wires, one for power and one for ground, going from it to the control box.  For large-scale controllers in factories, this was expensive and untenable to maintain~\cite{Thomesse2005}.  To solve this, practitioners developed fieldbus networks to handle their input and output.  The earliest was CAMAC which was used in the 1960s for nuclear instrumentation~\cite{Thomesse2005}.  Fieldbus networks provided benefits additional to wire cleanup.  They also enable distributing control information to many machines at once on a single network, monitoring the state of the factory floor from the design or business sections of the factory, and hierarchical control, such as described by the US National Bureau of Standards Automated Manufacturing Research Facility (AMRF) control hierarchy, where more abstract elements of an industrial plants command more concrete ones.  In particular, it describes a facility, consisting of production management and manufacturing engineering, commanding the shop, or factory floor, which in turn commands individual manufacturing cells, which then command the workstations within, and then at those workstations, operators command equipment~\cite{McLean1987}.  This all can be seen in Figure~\ref{fig:amrf_control_hierarchy}.  The control tooling discussed in this thesis is concerned with the business at the Equipment and Workstation level.

@@

This thesis builds on this history to improve the process of commissioning control systems for mechatronic systems by reasoning about how to abstract the construction of behavior and components in computer control systems into layers, synthesizing a novel {\it mechatronic tech stack\/} theory, not unlike the AMRF control hierarchy. It will demonstrate in four case studies how this theory can be applied in mechatronic systems. The tools used in these cases will be categorized into layers of the mechatronic tech stack.  The analysis shows that, while different, these cases all fit into the same structure.

@@

This section constructs the backbone of this thesis, namely the novel theory of a mechatronic tech stack. First, there will be a review of tech stacks and {\it abstraction layers\/} as they've been applied in the IT sector. Then there will be a proposal of a mechatronic tech stack determined by a standards-based set of principles.

@@

Tools in each of these layers are myriad and to highlight that, proceeding in Figure~\ref{fig:web_stack} is a list of a small sample. Front-end tools like jQuery~\cite{jQuery} have been proven from many years of use, React~\cite{React} is very powerful and can be used in non-web contexts, and even plain HTML without an additional framework can be good on account of its simplicity. Back-end tools can allow a developer to select their favorite programming languages like Ruby or Python, or one that is specialized for web back-ends like PHP, or even JavaScript like the front end uses. Servers are often integrated into the back-end framework, but if not, specialized servers like Apache~\cite{ApacheServer} or nginx~\cite{nginx} are popular. Databases are dominated by reliable and proven relational, SQL-based offerings like MySQL or PostgreSQL, but alternatives like NoSQL or Redis~\cite{redis} have become popular in the past years.

@@

The model itself consists of layered abstractions over physical communication media, as seen in Figure~\ref{fig:osi_model}. The lowest OSI layer, Physical, concerns itself with the physical phenomena of communication. In the case of CAN, this means specifying voltage levels, terminating resistors, twisted pairs, grounding rules, and more. The highest level, Application, provides networking facilities without need for understanding \textit{how} the network connection is made. The Application layer includes, for example, identifying communication partners and negotiating security protocols. Well-known protocols like HTTP, SSH, and FTP exist in the Application layer. The intermediate layers provide increasing levels of abstraction over the physical layer, defining things like the structure of data packets, how to route messages, how to route them efficiently, and more.

@@

When these tokens are combined into phrases, that's syntax. Syntax is the structure into which meaning is applied. The \verb|printf| example is syntactically parsed by C compilers as in Figure~\ref{fig:printf_syntax}. The syntax tells that the postfix expression represented by the identifier \verb|printf| has the argument list of \verb|"I have %u cows\n"| (a string literal) and \verb|num_cows| (an identifier). What any of this means is the domain of semantics.

@@

This is enough to propose a mechatronic tech stack. The mechatronic stack as outlined sits on top of signaling with sensors and transducers. Those are interfaced by IO systems that connect to computer systems which run middleware upon which a control program depends. The design is so that functionality in the control layer should be preserved if components in the lower layers are changed, perhaps with some implementation effort, as was the case for IP over CAN.  This is illustrated as a stack diagram in Figure~\ref{fig:mechatronic_tech_stack}.

@@

Following will be four case studies providing concrete illustrations of how this mechatronic tech stack model can be applied. These will mostly be descriptive case studies that apply this theory for retrospective research, but one case uses this model prescriptively.

@@

Dolores is a full-scale test bed in the Aalto University Fluid Power Laboratory, seen in Figure~\ref{fig:dolores0} to study the application of electric hydrostatic actuation in an articulated-steering linkage for a front-loading mining vehicle~\cite{Hermansson2021, Helduser1999, Dudzinski1989}. Its control system commissioning was completed in 2021, but in 2023, an industry partner in the experiment requested hardware changes that demanded a substantial re-architecture of the control system. This case study will discuss the mechatronic tech stacks behind the old and the new systems.

@@

Together, the stack for the old system was CANopen-based IO with a DAQ supplementing simple analog and digital signals, a workstation computer, Windows middleware with drivers and libraries to support communication with the CAN bus and DAQ, and a controller based on Matlab and Simulink. In the new system, the stack was a mix of EtherCAT and CANopen IO, a Beckhoff Industrial PC in concert with a workstation, PLC middleware with an application to initialize the CAN network, and a controller written in PLC languages with TwinCAT.\ This is illustrated in Figure~\ref{fig:mechatronic_tech_stack_dolores}.

@@

A research group at Okayama University of Science has, over the past years, developed a novel soft, pneumatic actuator that extends when supplied with pressure called an extension-type flexible pneumatic actuator (EFPA)~\cite{Noritsugu2005}. This section will study the tech stacks for three applications of this actuator, a six-legged mobile robot for providing the elderly with an interactive physical activity~\cite{Hase2020}, a pipe-traversal robot for inspecting and repairing water pipe networks~\cite{Shinohara2020}, and a lightweight servo valve based on bending hoses to support EFPA applications~\cite{Kobayashi2020}. All of these can be seen in Figure~\ref{fig:soft_pneumatic_actuators_demo}. Note that because this is a case study on work I did not participate in, some details are not documented and must be inferred.

@@

The system is a chair atop a Stewart platform whose linear actuators are single-acting pneumatic-muscle cylinders, seen in Figure~\ref{fig:stewart_platform_demo}~\cite{Stewart1965, Caldwell1995}. The valve manifolds that were already made for this had one valve to connect the pneumatic muscle to an external pressure source and another to vent the muscle to atmosphere. When both valves are closed, the cylinder brakes. There are six instances of this circuit, connected in series to a common pressure source. Each cylinder also has a position sensor exposed as a voltage signal from a potentiometer.

@@

For the control layer, I wrote the code in the Sublime Text text editor~\cite{SublimeText} and configured it to automatically format the source code according to standards for the Rust language using the \verb|rustfmt| utility~\cite{rustfmt}.  Other text editors like Emacs, Vim, or VSCodium would have also been good choices, but Sublime is familiar.  This was done on a Windows workstation, but it was compiled and run on the Raspberry Pi Linux control computer.  To achieve this, I checked the changes on the workstation into a \verb|git| version control repository and pushed them to the control computer\footnote{A small issue that comes up is that \verb|git| generally doesn't allow users to push changes to a branch of a repository that the local user has checked out, which is the case here~\cite[§receive.denyCurrentBranch]{GitConfig}.  This is to prevent data loss.  To enable this, \verb|git| can be configured to update the local branch only if there is no possibility of data loss by executing \verb|git config receive.denyCurrentBranch updateInstead|.}.  Then, though a remote terminal, I executed commands to compile and run the program.  This became quite annoying with how many steps there are, so I use the utility \verb|cargo-watch| to automatically execute the compile and run commands on the control computer when there are changes\footnote{The command was similar to \verb|cargo watch -x 'run -- eth0'|, which compiles and executes the program with the argument \verb|eth0|, which is the network interface the EtherCAT connection is on.}~\cite{CargoWatch}.  This still is somewhat tedious and so requires more development.  It also would have been mitigated by developing in some kind of Linux environment or by using a containerized environment like would be provided by Docker or Vagrant, which would have allowed running some tests on the development machine~\cite{DockerDev,VagrantDev}.  Most of the middleware was prepared with the Raspberry Pi OS image that was installed while commissioning the control computer, but there was some additional configuration, e.g.\ to install \verb|git|, that was done with the OS's \verb|apt| package management utility~\cite{apt}.  As well, library dependencies for the control program like Ethercrab were managed with the Rust \verb|cargo| utility~\cite{Cargo}.  The computer, once commissioned, did not require any development.  The IO layer's development was aided by a four-channel oscilloscope and a multimeter, seen in Figure~\ref{fig:stewart_platform_demo}, to validate electrical connections and ensure signal were propagating as expected.  It was also interesting to observe the timing jitter of the program, which was generally very small, on the order of microseconds.

@@

The Solid Mechanics Laboratory at Aalto University conducts experiments to identify the properties of engineering materials under load particularly studying phenomena ranging from the micron to the meter scale.  As is typical for laboratories studying solid mechanics, they do this by subjecting test samples to a load schedule using a servo-controlled hydraulic cylinder as seen in Figure~\ref{fig:solid_mechanics_test_rig}.  They have several test beds, each with an instance of a common control system.  This had been controlled by a workstation-commanded DAQ that communicated with a motor controller and a strain gauge amplifier, but new experiments and the desire for faster, better research cycles called for improvements to the control system.  The following discussion is based heavily on an interview I took with the lab's head Dr.~Pauli Lehto.

@@

The control layer was built, like the lower layers, with NI tooling.  All of it was built with LabView, with the CPU-based controls using the LabView Real-Time Module~\cite{NIRealTimeModule} and the FPGA controls built with the LabView FPGA Module~\cite{NIFPGAModule}.  It implemented a PID controller, though the derivative term was in practice usually omitted~\cite{Schad2011,Ali2010}.  To monitor the condition of long-running tests and for data acquisition, there are programs running on a desktop workstation, seen in Figure~\ref{fig:solid_mechanics_control_monitor}, that observe the state of the control system.  They also developed a custom scripting language that runs on the workstation to define the high-level testing schedule, separate from the low-level control logic.

@@

The Aalto Solid Mechanics Laboratory's new control system provides a compelling case study on how control components may be selected when there are requirements placed on the control system beyond easy development.  It led to a need for high-speed IO and computation and a computer with multiple computation units with a CPU and an FPGA.\  Its tech stack is illustrated in Figure~\ref{fig:mechatronic_tech_stack_solid_mechanics}.  The system as constructed has been able to meet the performance requirements and execute useful tests.  Development is on-going.

@@

This thesis has introduced a theory of a mechatronic tech stack which describes function and tooling for mechatronic systems, namely as discrete layers of control, middleware, computer, and IO.\  This theory is motivated by analogy to the web stack, used in IT software development, and to the OSI networking model, used to describe networked computer systems.  The OSI model also provided a means of discriminating the layers that make up the model, importantly by finding contrasts in the morphology, syntax, and semantics of the behavior of components of the control system.

@@

This theory can also be extended.  Throughout this work, it has been called a \textit{mechatronic} tech stack, but truly it's a control system tech stack.  To make it a true mechatronic tech stack, it should include tooling for mechanical and electronic systems.  A mechanical section of the stack would grow from the bottom of the stack shown in Figure~\ref{fig:mechatronic_tech_stack}, not colliding with the current stack, but an electrical component would be more complicated, since it is needed by both the mechanical stack for transducers and sensors, and by the control stack since the computer \textit{is} an electrical component, and the IO layer often needs some electrical finesse, as noted in §\ref{sec:stewart_platform}.  It could also be extended in the software realm; the production of graphical monitoring systems and tele-operation programs generally require substantial non-control software which should be described by different layers, perhaps forking away from the middleware layer to have a monitoring layer branch.

@@

This thesis presented four case studies to demonstrate the utility of the novel theory of a mechatronic tech stack. These cases were diverse in their applications and methods and by viewing them through the lens of the tech stack, new knowledge of how mechatronic systems can be commissioned was gained. Also, by studying several stacks, I have formed a pool of tools to draw from. The mechatronic tech stack can be re-rendered to reflect this, as here in Figure~\ref{fig:mechatronic_tech_stack_filled}.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{assets/mechatronic_tech_stack_filled}
  \caption{The mechatronic tech stack for all of the case studies in one, with tools for each project highlighted in different colors.  Not only does the mechatronic tech stack present layers of abstraction, it also provides an idea of what tools to choose, similar to the web stack in Figure~\ref{fig:web_stack}.}\label{fig:mechatronic_tech_stack_filled}
\end{figure}
